\steppingstone{The Dense Grid}
\label{stone:densegrid}

The Fibonacci Spine solved one important problem.

Repeated multiplication by \(\varphi\) no longer requires repeated arithmetic. It has become movement along an exact table.

That is wonderfully efficient.

Unfortunately, most PhiBase quantities are not powers of \(\varphi\).

They live between the rungs of the Spine.

If the Spine is one exact path through PhiBase, what do we do with everything else?

We organize it.

Suppose we choose a coefficient bound \(B\). Build every PhiBase quantity

\[
P(n,p)=n+p\varphi
\]

satisfying

\[
|n|\le B,
\qquad
|p|\le B.
\]

Keep only the positive values. Now sort them from the smallest to the largest.

Nothing has been approximated.

Nothing has been changed.

We have simply placed every exact quantity into its natural order.

That ordered collection is the \emph{Dense Grid}.

\begin{table}[htbp]
\centering
\caption{A small Dense Grid. Every entry is exact. The decimal value is shown only to illustrate the ordering.}
\label{tab:densegrid}
\begin{tabular}{rcc}
\hline
Rank & PhiBase & Decimal\\
\hline
1 & $P(5,-3)$ & 0.145898\\
2 & $P(2,-1)$ & 0.381966\\
3 & $P(-1,1)$ & 0.618034\\
4 & $P(1,0)$ & 1.000000\\
5 & $P(0,1)$ & 1.618034\\
6 & $P(1,1)$ & 2.618034\\
7 & $P(2,1)$ & 4.236068\\
8 & $P(2,2)$ & 5.236068\\
9 & $P(2,3)$ & 6.854102\\
\hline
\end{tabular}
\end{table}

The Spine passes through the Grid, but it does not define the Grid. Most entries are ordinary PhiBase quantities that simply happen to lie between neighbouring powers of \(\varphi\).

The important idea is not the table itself.

It is that the ordering has already been done.

If two PhiBase quantities must be compared, the answer is already present.

If a nearby value is needed, its neighbours are already known.

The arithmetic has become memory.

The Spine remembers one exact family.

The Dense Grid remembers the landscape around it.

Together they remove a great deal of repeated work without sacrificing exactness.

\section*{Builder's Notebook}

Choose

\[
B=2.
\]

Construct every positive PhiBase quantity satisfying

\[
|n|\le2,
\qquad
|p|\le2.
\]

Sort them into ascending order.

Mark the entries that belong to the Fibonacci Spine.

Notice how the remaining entries fill the spaces between the Spine values.

\section*{Looking Ahead}

The Spine and the Dense Grid now preserve everything we have learned.

The next question is no longer mathematical.

How can those two tables become a practical computational tool?
