\steppingstone{The PhiBase Slide Rule}
\label{stone:sliderule}

The Fibonacci Spine answered one question.

Repeated multiplication by \(\varphi\) no longer requires repeated arithmetic. Once the Spine has been built, moving one row upward multiplies by \(\varphi\). Moving one row downward divides by \(\varphi\).

That is wonderfully efficient.

Unfortunately, most PhiBase quantities are not powers of \(\varphi\).

They live between the rungs of the Spine.

The Dense Grid answered that question by placing every exact PhiBase quantity into its natural order.

Now another question appears.

Once the Spine and the Dense Grid have been built, why should they ever be built again?

A builder already knows the answer.

A cabinet maker keeps templates.

A machinist keeps jigs.

An architect keeps drawings.

Successful work is preserved so that it never has to be rediscovered.

The PhiBase Slide Rule follows exactly the same idea.

The Spine remembers every exact power of \(\varphi\).

The Dense Grid remembers the exact quantities that lie between those powers.

Together they preserve the arithmetic.

Nothing has been approximated.

Nothing has been simplified.

The difficult work has simply been moved into preparation.

Once the tables exist, they become tools.

Suppose we wish to multiply

\[
\varphi^{7}
\]

by

\[
\varphi^{4}.
\]

The Spine answers immediately.

Move four rows.

The result is

\[
\varphi^{11}.
\]

No multiplication has been performed.

The arithmetic became navigation.

Now suppose we wish to compare

\[
P(8,5)
\]

with

\[
P(11,3).
\]

The Dense Grid already contains both quantities.

Their order has already been determined.

The comparison becomes a lookup instead of a fresh calculation.

The mathematics has not changed.

Only the repeated work has disappeared.

That is exactly what engineers have always tried to accomplish.

A slide rule never changed arithmetic.

It remembered relationships that had already been established.

The PhiBase Slide Rule does the same thing.

The only difference is that the remembered relationships are exact.

The Spine remembers one exact family.

The Dense Grid remembers the surrounding landscape.

Together they become a permanent computational tool.

\section*{Builder's Notebook}

Choose two neighbouring entries from the Spine.

Multiply the first by \(\varphi\).

Verify that the answer is the second.

Now choose two neighbouring entries from the Dense Grid.

Compare them first by their decimal values and then by their positions in the Grid.

Both methods give the same answer.

One simply remembers what the other computes.

\section*{Looking Ahead}

The Slide Rule preserves arithmetic.

The next question is whether the same ideas can preserve geometry.

Can an exact construction in space be remembered just as naturally as an exact number?
