\steppingstone{Building Space}
\label{stone:space}

When we were building on a line, life was simple. A single PhiBase quantity described every length because there was only one direction in which we could continue. The moment we stepped onto a flat surface, that simplicity disappeared. We could move in many directions, and our arithmetic had to learn how to remember places instead of lengths.

The ordinary mathematical answer was straightforward. Two independent directions are enough to describe every point on a plane, and our turtle drawings have quietly been using exactly that idea from the beginning. Every point in every figure is stored as two exact PhiBase quantities, one measured along the \(0^{\circ}\) direction and the other along the \(72^{\circ}\) direction. The mathematics remains exact until the final picture is drawn.

If our only goal were to describe where a point is, we could stop there.

Building asks a different question.

Imagine a robot laying golden triangles across a large floor. For a while one direction is the natural direction to follow. Then the robot reaches the edge of a pentagonal region and another direction immediately becomes the better choice. The point has not changed, but the construction has. The robot now has a different floor beneath its feet.

That small observation changes everything.

We are no longer asking how many directions are sufficient to describe a point.

We are asking how many useful directions should be remembered so that construction can continue naturally when the local floor changes.

Those are different questions.

Mathematics answers the first.

Construction answers the second.

That is why the previous stepping stone deliberately remembered five planar directions even though only two were mathematically necessary. The additional directions did not make the representation more accurate. They preserved useful possibilities. When the preferred direction changed, the next useful direction was already available. Nothing had to be rebuilt.

Space presents exactly the same choice.

Three independent directions are enough to describe every point in ordinary space. If our purpose were only representation, the familiar Cartesian coordinates

\[
(x,y,z)
\]

would already have solved the problem.

Our purpose is different.

We want a representation that grows naturally with the construction itself.

Instead of remembering only three directions, PhiBase remembers six construction floors. Each floor corresponds to one family of parallel dodecahedral faces. At any moment one floor is beneath the builder, while the other five remain immediately available should the construction cross onto another face.

A point is therefore described by six construction counts,

\[
(a,b,c,d,e,f).
\]

These are not six independent dimensions.

They are six remembered construction directions.

The point itself still occupies one place in ordinary three-dimensional space. The additional directions simply preserve the useful alternatives that make construction continuous instead of awkward.

The six construction floors are represented by the vectors

\[
\begin{aligned}
\mathbf{F}_1 &= (\varphi,\,1,\,0), &
\mathbf{F}_2 &= (\varphi,\,-1,\,0), \\
\mathbf{F}_3 &= (1,\,0,\,\varphi), &
\mathbf{F}_4 &= (-1,\,0,\,\varphi), \\
\mathbf{F}_5 &= (0,\,\varphi,\,1), &
\mathbf{F}_6 &= (0,\,\varphi,\,-1).
\end{aligned}
\]

Beginning at the origin, the completed construction is

\[
a\mathbf{F}_1 + b\mathbf{F}_2 + c\mathbf{F}_3 + d\mathbf{F}_4 + e\mathbf{F}_5 + f\mathbf{F}_6.
\]

Collecting the components gives the ordinary Cartesian position,

\[
\begin{aligned}
x &= \varphi(a+b) + (c-d), \\
y &= (a-b) + \varphi(e+f), \\
z &= \varphi(c+d) + (e-f).
\end{aligned}
\]

Nothing has been lost.

The Cartesian coordinates tell us where the point is.

The six-floor description remembers how the construction reached that point and which neighbouring floors are immediately available when the construction continues.

One description serves the surveyor.

The other serves the builder.

Both describe exactly the same point.

\section*{Builder's Notebook}

Begin with the construction

\[
(1,0,0,0,0,0).
\]

Convert it into Cartesian coordinates.

Now imagine that the next step carries the construction onto a neighbouring dodecahedral face.

Which of the six construction floors becomes the natural one to continue from?

The point has not changed.

Only the floor beneath the builder has.

\section*{Looking Ahead}

The builder's description and the surveyor's description describe the same point in different ways.

The next stepping stone shows how to move from one description to the other and back again without losing a single piece of information.